---
title: "Voting Data Analysis"
author: "Jenn Schilling"
date: "8/22/2020"
output: html_document
---

# Plan

Explore election/voting data available from the U.S. Election Assistance Commission to show the percentage of voters who vote by mail, have mail-in ballots rejected, and the reasons for mail in ballot rejection. Examine data from 2016 and 2018. Compare these years to data collected by NPR in 2020. 

## Data Points Needed

### U.S. Election Assistance Commission 
* FIPSCode (Jurisdiction identification code)
* State
* Jurisdiction Name
* Number of registered, eligible voters (A1a)
* Number of active, registered, eligible voters (A3a)
* Number of domestic civilian absentee/mail-in ballots transmitted to voters (C1a)
* Number of absentee/mail-in ballots returned for counting (C1b)
* Number of absentee/mail-in ballots counted (2016 C4a, 2018 C3a)
* Number of absentee/mail-in ballots rejected (2016 C4b, 2018 C4a)  
* Number rejected by reason 
  + Missed deadline (2016 C5a, 2018 C4b)  
  + No signature (2016 C5b, 2018 C4c)  
  + No witness signature (2016 C5c, 2018 C4d)  
  + Non-matching signature (2016 C5d, 2018 C4e)  
  + No election official's signature (2016 C5e, 2018 C4f)  
  + Unofficial envelope (2016 C5f, 2018 C4g)  
  + Ballot missing from envelope (2016 C5g, 2018 C4h)  
  + Envelope not sealed (2016 C5h, 2018 C4i)  
  + No resident address on envelope (2016 C5i, 2018 C4j)  
  + Multiple ballots in one envelope (2016 C5j, 2018 C4k)  
  + Voter deceased (2016 C5k, 2018 C4l)  
  + Voted in person (2016 C5l, 2018 C4m)  
  + First-time voter without proper identification (2016 C5m, 2018 C4n)  
  + No ballot application on record (2016 C5n, 2018 C4o)  
  + Other (2016 C5o-C5v, 2018 C4o-C4r)  
* Total votes cast (2016 D1a, 2018 D1a)

# Setup  

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

# Libaries
library(tidyverse)

# Data

# Election Administration and Voting Survey Data - 2016
eavs_2016 <- read.csv("EAVS 2016 Final Data for Public Release v.3.csv", 
                      stringsAsFactors = FALSE)

# Election Administration and Voting Survey Data - 2018
eavs_2018 <- read.csv("EAVS_2018_for_Public_Release_Updates3.csv", 
                      stringsAsFactors = FALSE)

# NPR Article - 2020
# Will need to learn how to get the table via webscraping

```

# Data Processing 

```{r processing}

# Get variables of interest
# Fill in NAs
eavs_2016_p <- eavs_2016 %>%
  select(FIPSCode,
         State,
         JurisdictionName,
         A1a,
         A3a,
         C1a,
         C1b,
         C4a,
         C4b,
         C5a,
         C5b,
         C5c,
         C5d,
         C5e,
         C5f,
         C5g,
         C5h,
         C5i,
         C5j,
         C5k,
         C5l,
         C5m,
         C5n,
         D1a) %>%
  pivot_longer(A1a:D1a, names_to = "names", values_to = "value") %>%
  mutate(value = ifelse(str_detect(value, "Not Applicable"), NA, value)) %>%
  pivot_wider(names_from = "names", values_from = "value")


```



Note: another interesting set of data points to examine would be how many voters are removed in different categories by state (A11 questions)